### 2021-07-14 src/paper/literature_review.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

\citet{Bicher2021} simulate the entire Austrian population. The use data from the first
wave (February 21 to April 9, 2020) to calibrate ther model and predict the effect of
different NPIs and contact tracing policies until November 2020. They the same data
provided by \citet{Mossong2008} as we use to calibrate contact networks for households,
workplaces and schools. The model focuses on analyzing the effect of different contact
tracing strategies and not on modelling enacted Austrian policies over a long period
of time.

@@

Our model combines elements from the above models and adds other ones. To the best of
our knowledge, our model is the only one with the following features:

1. The free model parameters have been estimated with the method of simulated moments
\citep{McFadden1989}. Despite having few free parameters our model does en excellent
job in explaining observed case numbers and the spread of the B.1.1.7 mutation over
more than nine months of data.

2. We have an extremely fine grained of schools and preschools.
We can thus easily model all NPIs that have been implemented in Germany in the past 9
months. This includes complete school closures, phases where only those students whose
parents could not find any private childcare arrangement could attend, split class
approaches for some or all age groups and combinations thereof. Moreover, we can
account for additional hygiene measures whose effect is estimated inside the model.

3. we model the evolution of the pandemic and all enacted policies over since the
start of the second wave. Since the vast majority of cases has occurred in that time
period and we also model unobserved infections our simulations take into account that
many people are already immune because they are recovered from an infection and that
this immunity is not spread randomly across the population.

4. We have an extremely detailed model of PCR and rapid tests with a share of detected
cases that varies over time and across age groups. We fit estimates of the share of
known cases from other studies for time periods where those are available but also
estimate a plausible share of known cases after vaccinations become widespread and
methodologies that rely on a case fatality ratio to identify the share of known
cases are not applicable anymore.

>>>>

\citet{Bicher2021} simulate the entire Austrian population. They use data from the first
wave (February 21 to April 9, 2020) to calibrate their model and predict the effect of
different NPIs and contact tracing policies until November 2020. They use the same data
provided by \citet{Mossong2008} as we to calibrate contact networks for households,
workplaces and schools. The model focuses on analyzing the effect of different contact
tracing strategies and not on modelling enacted Austrian policies over a long period
of time.

@@

Our model combines elements from the above models and adds several others. To the best of
our knowledge, our model is the only one with the following features:

1. The free model parameters have been estimated with the method of simulated moments
   \citep{McFadden1989}. Despite having few free parameters our model does an excellent
   job in explaining observed case numbers and the spread of the B.1.1.7 mutation over
   more than nine months of data.

2. We have an extremely fine grained \replaced[id=K]{representation}{} of schools and
   preschools. We can thus easily model all \replaced[id=K]{schooling policies}{NPIs}
   that have been implemented in Germany in the past nine months. This includes complete
   school closures, phases where only those students whose parents could not find any
   private childcare arrangement could attend, split class approaches for some or all age
   groups and combinations thereof. Moreover, we can account for additional hygiene
   measures whose effect is estimated inside the model.

3. We model the evolution of the pandemic and all enacted policies since the start of the
   second wave. Since the vast majority of cases has occurred in that time period and we
   also model unobserved infections our simulations take into account that many people
   are already immune because they have recovered from an infection and that this
   immunity is not spread randomly across the population.

4. We have an extremely detailed model of PCR and rapid tests with a share of detected
   cases that varies over time and across age groups. We fit estimates of the share of
   known cases from other studies for time periods where those are
   available\comment[id=K]{Citation here?} but also estimate a plausible share of known
   cases after vaccinations become widespread and methodologies that rely on the case
   fatality ratio to identify the share of known cases are not applicable anymore.

 \comment[id=K]{Four is of course plenty but what about weekly recurring contacts or using the google mobility data? -> maybe add these to point 2?}

### 2021-07-14 src/paper/model/course_of_disease.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

The disease progression in the model is fairly standard. It is depicted in
Figure~\ref{fig:model_disease_progression} and the values and source of the relevant
parameters are describes in Section~\ref{sub:data_course_of_disease}.

@@

We allow the progression of the disease to be stochastic in two ways: Firstly, state
changes only occur with a certain probability (e.g. only a fraction of infected
individuals develops symptoms). Secondly, the number of periods for which an individual
remains in a state is drawn randomly. The parameters that govern these processes are
taken from the literature \footnote{ Detailed information on the calibration of the
disease parameters is available as part of our
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.}.\comment[id=J]{Replace this by a link to the relevant table with
parameters in the data appendix, once we have that table.}

>>>>

The disease progression in the model is fairly standard. It is depicted in
Figure~\ref{fig:model_disease_progression} and the values and source of the relevant
parameters are described in Section~\ref{sub:data_course_of_disease}.

@@

We allow the progression of the disease to be stochastic in two ways: Firstly, state
changes only occur with a certain probability (e.g. only a fraction of infected
individuals develops symptoms). Secondly, the number of periods for which an individual
remains in a state is drawn randomly. The parameters that govern these processes are
taken from the literature and detailed in Section~\ref{sub:data_course_of_disease}. For
an overview of our disease progression parameters see
Table~\ref{tab:disease_and_vacc_model}.\comment[id=K]{@Janos: Is this what you wanted
("Replace this by a link to the relevant table with parameters in the data appendix, once
we have that table")}

### 2021-07-14 src/paper/model/matching.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

The empirical data described above only allows to estimate the number of contacts each
person has. In order to simulate transmissions of Covid-19, the numbers of contacts has
to be translated into actual meetings between people. This is achieved by a matching
algorithm:

@@

%% Recurrent contacts ============================================================
Recurrent contacts are described by two components: 1) A set of time invariant groups,
such as school classes or groups of co-workers. Those groups are generated once at the
beginning of the simulation. The groups can be sampled from empirical data or created by
randomly matching simulated individuals into groups. 2) A deterministic or random
function that takes the value 0 (non-participating) and 1 (participating) and can depend
on the weekday, date and health state. This can be used to model vacations, weekends or
symptomatic people who stay home (see section \ref{sec:policies} for details).

@@

where $beta_c$ denotes the base infection probability of contact type $c$,
$s_{c, t}$ is a seasonality factor between zero and one that depends on the contact
type $c$ and time $t$ (see Equation~\ref{eq:seasonality}), $\sigma_v$ is the
infectiousness factor of virus variant $v$ and $\zeta_a$ is an age dependent
susceptibility factor.

@@

The matching of non-recurrent contact types is more difficult because the contact
network is resampled randomly every day. Moreover, it needs to allow for
assortative matching. In our application, all
random contacts are assortative with respect to age group $a$ (it is usually more likely
to meet
people from the same age group) and county (it is more likely to meet people from the
same county) but in principle any set of discrete variables can be used. This set of
variables that influence matching probabilities introduce a discrete partition of the
population into groups.

@@

\begin{equation}
    \label{eq:combined_matching_probabilities}
    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual probability}}
\end{equation}

@@

\begin{align}
    K_{z+1,i,c} = K_{z, i, c} - 1
    K_{z+1, j, c} = K_{z, j, c} - 1
\end{align}

>>>>

The empirical data described above only allows to estimate the number of contacts each
person has. In order to simulate transmissions of Covid-19, the numbers of contacts have
to be translated into actual meetings between people. This is achieved by a matching
algorithm:

@@

%% Recurrent contacts ============================================================
Recurrent contacts are described by two components: 1) A set of time invariant groups,
such as school classes or groups of co-workers. Those groups are generated once at the
beginning of the simulation. The groups can be sampled from empirical data or created by
randomly matching simulated individuals into groups. 2) A deterministic or random
function that takes the value 0 (non-participating) and 1 (participating) and can depend
on the weekday, date and health state.\comment[id=K]{Actually the full states and params
enter into the contact functions} This can be used to model things like vacations,
weekends or symptomatic people who stay home (see section \ref{sec:policies} for
details).

@@

where $\beta_c$ denotes the base infection probability of contact type $c$,
$s_{c, t}$ is a seasonality factor between zero and one that depends on the contact
type $c$ and time $t$ (see Equation~\ref{eq:seasonality}), $\sigma_v$ is the
infectiousness factor of virus variant $v$ and $\zeta_a$ is an age dependent
susceptibility factor.

@@

The matching of non-recurrent contact types is more difficult because the contact network
is resampled randomly every day. Moreover, it needs to allow for assortative matching. In
our application, all random contacts are assortative with respect to age group $a$ (it is
usually more likely to meet people from the same age group) and county (it is more likely
to meet people from the same county) but in principle any set of discrete variables can
be used. This set of variables that influence matching probabilities
\replaced[id=K]{induce}{introduce} a discrete partition of the population into groups.

@@

\begin{equation}
    \label{eq:combined_matching_probabilities}
    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual Probability}}
\end{equation}

@@

\begin{align}
    K_{z+1,i,c} &= K_{z, i, c} - 1 \\
    K_{z+1, j, c} &= K_{z, j, c} - 1
\end{align}

### 2021-07-14 src/paper/model/policies.tex
Add first version of estimated parameters table.
Fix typos, make suggestions and add comments to sections B.1 to B.7.

Our model makes it very easy to model a wide range of NPIs, either in isolation or
simultaneously. This is important for two reasons: Firstly, it allows to predict and
quantify the effect of novel NPIs. Secondly, it allows to model the actually implemented
policy environment in great detail, which is necessary to use use the full time series
of infections and fatality rates to estimate the model parameters.\footnote{
See \citet{Avery2020} for an explanation why it can be harmful to use too long time
series to estimate simple SEIR type models.}

@@

The most complex policies are typically found in the education sector. Since the
beginning of 2021 schools have switched back and fourth between full closures,
split class approaches with alternating schedules for some or all age groups and
reopening while maintaining hygiene measures. On top of that there are different
policies for allowing young students whose parents work full time to attend school
even on days where they normally would not. For details on how we calibrate these
policies see Section~\ref{subsec:policies_seasonality}.

Importantly, policies can depend on the health states of participating individuals.
This allows to quarantine entire school classes if one student tested positive or
to implement official or private contact tracing.

For some policies the exact effect on each contact type is not easy to determine. If
this refers to a policy has been active during the estimation period, it is possible to
estimate such parameters by fitting the model to time series data of infection rates.
This is only possible if the policy was not active during the whole estimation period
and thus the infection probabilities can be identified separately. We do this to account
for hygiene measures at school and in the workplace that have been in effect since
November 2020.

Not all things that reduce contacts compared to the pre-pandemic level are driven
by NPIs. Therefore, we also model endogenous contact reductions that can depend on
the health state of individuals, known risk contacts or the local incidence of
infections. Examples are strong contact reductions for symptomatic individuals or those
who have a positive PCR or rapid tests or contact reductions when a houshold member
tested positive. The extent to which contacts are reduced can be calibrated from
surveys. For an application of our model showcasing private contact tracing in the
context of the Christmas holidays see \cite{Gabler2020}.

>>>>

 \comment[id=K]{I think we need to make sure we distinguish very clearly what is possible
 and what's implemented.}

Our model makes it very easy to model a wide range of NPIs, either in isolation or
simultaneously. This is important for two reasons: Firstly, it allows to predict and
quantify the effect of novel NPIs. Secondly, it allows to model the actually implemented
policy environment in great detail, which is necessary to use the full time series
of infections and fatality rates to estimate the model parameters.\footnote{
See \citet{Avery2020} for an explanation why it can be harmful to use too long time
series to estimate simple SEIR type models.}

@@

The most complex policies are typically found in the education sector. Since the
beginning of \replaced[id=K]{2020}{2021} schools have switched back and fourth between
full closures, split class approaches with alternating schedules for some or all age
groups and reopening while maintaining hygiene measures. On top of that there are
different policies for allowing young students whose parents work full time to attend
school even on days where they normally would not. For details on how we calibrate these
policies see Section~\ref{subsec:policies}.\comment[id=K]{Did I write enough on the school policies? (No sources or details so far...)}

Importantly, policies can depend on the health states of participating individuals.
\replaced[id=K]{Thus, it would be possible}{This allows} to quarantine entire school
classes if one student tested positive \replaced[id=K]{and many other forms of contact
tracing}{or to implement official or private contact tracing}.

For some policies the exact effect on each contact type is not easy to determine. If this
refers to a policy that has been active during the estimation period, it is possible to
estimate such parameters by fitting the model to time series data of infection rates.
This is only possible if the policy was not active during the whole estimation period and
thus the infection probabilities can be identified separately. We do this to account for
hygiene measures at school and in the workplace that have been in effect since November
2020.\comment[id=K]{I don't understand this paragraph.}

Not all things that reduce contacts compared to the pre-pandemic level are driven by
NPIs. Therefore, we also model endogenous contact reductions that can depend on the
health state of individuals \replaced[id=K]{}{, known risk contacts} or the local
incidence. Examples in our model are contact reductions for symptomatic individuals or
those who have a positive PCR or rapid tests \replaced[id=K]{or contact reductions when a
household member tested positive}. The extent to which contacts are reduced can be
calibrated from surveys. For an application of our model showcasing private contact
tracing in the context of the Christmas holidays see \cite{Gabler2020}.

### 2021-07-14 src/paper/model/summary.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

To predict and quantify the effects of a wide variety of fine-grained social distancing
policies, vaccinations and rapid testing, we propose a different model structure. Our
model inherits many features from prototypical agent-based simulation models but
replaces the contacts between moving particles by contacts between individuals who work,
go to school, live in a household and enjoy leisure activities.

@@

After rapid tests become available, people who work or go to school can receive rapid
tests there. Moreover, people can decide to make a rapid test if they develop symptoms,
have many planned contacts or observe cases in their contact network. People who have a
positive rapid test demand a confirmatory PCR test with a certain probability. Moreover,
PCR tests can be demanded because of symptoms or randomly.

@@

People who have symptoms, received a positive test, or had a risk contact can reduce
their number of contacts across all contact types endogenously. The extent to which this
is done is calibrated from survey data.

@@

The model has a very modular structure and can easily be extended to distinguish more
contact types, add more stages to the disease progression, implement new policies
or test demand models. The main bottleneck is not complexity or computational cost
but the availability of data to calibrate the additional model features.

>>>>

To predict and quantify the effects of a wide variety of fine-grained social distancing
policies, vaccinations and rapid testing, we \replaced[id=K]{use the following}{propose a
different} model structure. \replaced[id=K]{}{It inherits many features from prototypical
agent-based simulation models but replaces the contacts between moving particles by
contacts between individuals who work, go to school, live in a household and enjoy
leisure activities.}\comment[id=K]{This has been moved to the Literature Review, I
think.}

@@

After rapid tests become available, people who work or go to school can receive rapid
tests there. Moreover, people can decide to do a rapid test if they develop symptoms,
have many planned contacts\comment[id=K]{at the moment it's implemented as planing to
participate in a weekly leisure meeting. Should we be more explicit here?} or
\replaced[id=K]{have a sick or positively tested household member}{observe cases in their
contact network}. People who have a positive rapid test demand a confirmatory PCR test
with a certain probability. Moreover, PCR tests can be demanded because of symptoms or
randomly.

@@

People who have symptoms\replaced[id=K]{ or}{,} received a positive test,
\replaced[id=K]{}{or had a risk contact} can reduce their number of contacts across all
contact types endogenously. The extent to which this is done is calibrated from survey
data.

@@

The model has a very modular structure and can easily be extended to distinguish more
contact types, add more stages to the disease progression, implement new policies
or test demand models. The main bottleneck is not complexity or computational cost
but the availability of data to calibrate additional model features.

### 2021-07-14 src/paper/model/testing.tex
Fix typos, make suggestions and add comments to sections B.1 to B.7.

Having a realistic model of PCR and rapid tests is crucial for two reasons: Firstly,
only via a testing model the simulated infections from the model can be made comparable
to official case numbers. Secondly, individuals with undetected or not yet detected
infections are an important driver of the pandemic.

@@

Therefore, we aim for a simpler model of PCR and rapid tests that can be calibrated with
available data on test demand and availability and -- nevertheless -- can produce a share
of undetected cases that varies over time and across age groups and agrees with other
estimates over the time periods where they are available.

PCR tests are modeled since the beginning of the simulation and determine whether a
infections is officially recorded. Rapid tests are only added at the beginning of 2021.
Positive rapid tests do not enter official case numbers directly, but most people with a
positive rapid tests demand a confirmatory PCR test. However, positive rapid tests can
have a strong effect on the infection dynamics because they trigger contact reductions
and additional rapid tests.

@@

At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
of detected cases from \citet{Dunkelzifferradar2020} can no longer be used because it is
based on case fatality rates which drastically change due to vaccinations. Secondly,
rapid tests become available at a large scale.

@@

The second challenge is solved by mechanistic rapid test demand models for the
workplace, schools and by private individuals. The calibration of these models is
described in Section~\ref{subsec:rapid_test_demand}.
Figure~\ref{fig:antigen_tests_vaccinations} shows that the number of performed rapid
tests in the model fits the empirical data well (where empirical data is available).

In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
falsely negative. While the specificity of rapid tests is constant over time, their
sensitivity strongly depends on whether the tested individual is already infectious and
if so for how long he has been infectious. Before the onset of infectiousness the
sensitivity is very low (35\%). On the first day of infectiousness it is much higher
(88\%) but still lower than during the remaining infectious period (92\%). After
infectiousness stops, the sensitivity drops to 50 \%.

@@

While rapid tests do not directly enter official case numbers, many positively tested
individuals confirm their rapid test result with a PCR test.\comment[id=J]{Need to link
to data appendix once the number is there and maybe mention the number}. Importantly,
those PCR tests are made in addition to the tests that would have been done otherwise.
Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid tests on
the share of detected cases.\comment[id=K]{Talk here about reactions to tests and
symptoms with contact reductions. (If not here, find other place where to talk about
this in detail.)}

>>>>

 \comment[id=K]{Talk here about reactions to tests
and symptoms with contact reductions. (If not here, find other place where to talk about
this in detail.)}

Having a realistic model of PCR and rapid tests is crucial for two reasons: Firstly, only
via a testing model can the simulated infections from the model be made comparable to
official case numbers. Secondly, individuals with undetected or not yet detected
infections are an important driver of the pandemic: \textcolor{blue}{One of the features
that makes SARS-CoV-2 so hard to control is that it can be transmitted both by pre- and
asymptomatic individuals which makes it difficult for individuals to isolate when they
pose a risk to others. Our model captures this because developing symptoms or receiving a
positive test result leads many people to reduce their contacts, especially contacts
outside the own household.}\comment[id=K]{I think these multipliers $\tau$ are not
mentioned in the data section.}

@@

Therefore, we \replaced[id=K]{build}{aim for} a simpler model of PCR and rapid tests that
can be calibrated with available data on test demand and availability and -- nevertheless
-- can produce a share of undetected cases that varies over time and across age groups
and agrees with other estimates over the time periods where they are available.

PCR tests are modeled since the beginning of the simulation and determine whether a
infection is officially recorded. Rapid tests are only added at the beginning of 2021.
Positive rapid tests do not enter official case numbers directly, but most people with a
positive rapid tests demand a confirmatory PCR test. However, positive rapid tests can
have a strong effect on the infection dynamics because they trigger contact reductions
and additional rapid tests.

@@

At the beginning of 2021, two challenges arise: Firstly, the externally estimated share
of detected cases from Dunkelzifferradar project \citep{Dunkelzifferradar2020} can no
longer be used because it is based on the case fatality rate which changes drastically
due to vaccinations. Secondly, rapid tests become available at a large scale.

@@

The second challenge is solved by mechanistic rapid test demand models for the
workplace, schools and by private individuals. The calibration of these models is
described in Section~\ref{subsec:rapid_test_demand}.
Figure~\ref{fig:antigen_tests_vaccinations} shows that the number of performed rapid
tests in the model fits the empirical data well (where empirical data is available).
 \comment[id=K]{Maybe refer to the more detailed \ref{fig:share_rapid_test_last_week2}}

In contrast to PCR tests, rapid tests are not perfect and can be falsely positive or
falsely negative. While the specificity of rapid tests is constant, their sensitivity
strongly depends \replaced[id=K]{on the timing of the rapid test relative to the start of
infectiousness}{on whether the tested individual is already infectious and if so for how
long he has been infectious}. Before the onset of infectiousness the sensitivity is very
low (35\%). On the first day of infectiousness it is much higher (88\%) but still lower
than during the remaining infectious period (92\%). After infectiousness stops, the
sensitivity drops to 50 \%.

@@

While rapid tests do not directly enter official case numbers, 82\%
($\chi_{confirmation}$) of positively tested individuals seek a PCR test
\citep{Betsch2021}.\comment[id=J]{Need to link to data appendix once the number is there
and maybe mention the number}\comment[id=K]{Not mentioned in the data section.}
Importantly, those PCR tests are made in addition to the tests that would have been done
otherwise. Section~\ref{subsec:results_share_known_cases} discusses the effect of rapid
tests on the share of detected cases.
